---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(dplyr)
library(sf)
minimal_mesh <- silicate::minimal_mesh
sf::st_crs(minimal_mesh) <- NA
inlandwaters <- silicate::inlandwaters
sf::st_crs(inlandwaters) <- NA

```

# area

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R build status](https://github.com/hypertidy/area/workflows/R-CMD-check/badge.svg)](https://github.com/hypertidy/area/actions)
<!-- badges: end -->

The goal of area is to calculate areas, allow control over how that happens, and illustrate how it works. 

Very much work in progress. 

Please, do not use this for your land surveying contract. 

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("hypertidy/area")
```

## Straightforward tools

The area package is written to allow easy to use calculations for
developing packages. It's similar to the [traipse](https://CRAN.r-project.org/) package which was created to leverage the grouping mechanisms of the tidyverse for common tracking data calculations. 


```{r areas}
library(area)
polygon_area(cbind(c(0, 1, 1, 0, 0), 
                c(0, 0, 1, 1, 0)))

triangle_area(cbind(c(0, 1, 1, 1, 0, 0), 
               c(0, 0, 1, 1, 1, 0)))

```

There's an inbuilt data set with a path of Tasmania's outline. 

```{r tas}
index <- 4423 ## explain later
idx <- c(tas$S[1], tas$S[1:index,2], 1)
plot(tas$P[idx, ], type = "l", asp = 1)
polygon_area(tas$P[idx, ])/1e6

sf::st_area(sf::st_polygon(list(tas$P[idx, ])))/1e6
```


```{r area}
library(dplyr)
library(area)

calc_area <- function(df) {
 area <- df %>% mutate(row = row_number()) %>% 
  group_by(multipolygon_id, polygon_id, linestring_id) %>% 
   arrange(row) %>% 
  summarize(area = polygon_area(cbind(x, y))) %>% 
  ungroup() #%>% 
  area %>%  mutate(area = area, 
                   hole = c(-1, 1)[(pmax(c(1, diff(multipolygon_id)), c(1, diff(polygon_id))) == 1) + 1])
}
xsf <- silicate::inlandwaters
df <- sfheaders::sf_to_df(xsf)
(mmarea <- calc_area(df))
sum(mmarea$area * mmarea$hole)
sum(sf::st_area(xsf))






```

## Straightforward? 

Well - ha ha - we will need some ease of use here, and some speed ups - but we aren't restricted to one in-memory format, we have the area of every component part to use as we wish (the grouping tells us about holes and multipolygons), we can tell the orientation of the polygon or triangle and get its area, and when we have the right tools we'll only be using R and C++. Oh and we can calculate with triangles stored compactly.  


```{r triangles}
(a <- triangle_area(mm_tri$P[t(mm_tri$T), ]))
sum(a)

## so we don't double count the hole
print(sum(mmarea$area[mmarea$hole > 0]))
```

The key motivation here is *flexibility* and working with triangles. 

Performance is compelling, compare silicate::tri_area() to this C++ version

```{r tri-perf}
library(silicate)
tas_tri <- RTriangle::triangulate(tas)
trxy <- tas_tri$P[t(tas_tri$T), ]
dim(trxy)
rbenchmark::benchmark(R = {a <- tri_area(trxy)}, cpp = {b <- triangle_area(trxy)})
sum(abs(a - b))
mean(c(sum(a), sum(b)))
```

---

## Code of Conduct

Please note that the area project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
